---
title: 'Practical Machine Learning: Project Report'
author: "Alexander Zhou"
date: "12/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.path = "./figures/")
require(ggplot2)
require(caret)
require(corrplot)
require(Rtsne)
require(xgboost)
require(stats)
require(ggplot2)
```

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. The goal of this project is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants who were asked to perform barbell lifts correctly and incorrectly in 5 different ways, represented as classes in the training set. 

## Data Processing

```{r download}
if (!file.exists("pml-training.csv")) { 
  download.file(
    "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", 
    "pml-training.csv")
}
if (!file.exists("pml-testing.csv")) { 
  download.file(
    "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", 
    "pml-testing.csv")
}

train = read.csv("pml-training.csv")
test = read.csv("pml-testing.csv")

dim(train)
dim(test)
```
